\subsection{Hamiltonian}

For a generic continuous-time optimal control problem of the form
\begin{equation}
    \max_{\{u(t)\}} \int_0^\infty e^{-\rho t} I\bigl(x(t), u(t)\bigr)\,\mathrm{d}t
    \quad \text{subject to} \quad
    \dot{x}(t) = g\bigl(x(t), u(t)\bigr), \quad x(0) \text{ given},
\end{equation}
we define the \textbf{current-value Hamiltonian}
\begin{equation}
    \mathcal{H}(x, u, \lambda) = I(x, u) + \lambda\, g(x, u),
\end{equation}
where $\lambda(t)$ is the \emph{costate variable} (the continuous-time shadow value of the
state $x$). The necessary optimality conditions are
\begin{align}
    \mathcal{H}_u &= 0, \label{eq:foc_u}\\
    \dot{x} &= g(x, u), \label{eq:state}\\
    \dot{\lambda} &= \rho\lambda - \mathcal{H}_x. \label{eq:costate}
\end{align}
Together with the \textbf{transversality condition} (TVC)
\begin{equation}
    \lim_{t\to\infty} e^{-\rho t}\,\lambda(t)\,x(t) = 0,
\end{equation}
these conditions pin down the unique optimal path.

\paragraph{Household example.}
For a representative household with CRRA utility maximising
$\int_0^\infty e^{-\rho t} u(c)\,\mathrm{d}t$ subject to
$\dot{k} = rk + w - c$, the current-value Hamiltonian is
$\mathcal{H} = u(c) + \lambda(rk + w - c)$.
The FOCs yield $\lambda = u'(c)$ and $\dot{\lambda} = \lambda(\rho - r)$, which combine to
give the continuous-time \textbf{Euler equation}
\begin{equation}
    \frac{\dot{c}}{c} = \frac{1}{\sigma}(r - \rho).
\end{equation}
For the Ramsey model, replacing $r$ with the net marginal product of capital $f'(k)-\delta$
gives the canonical two-dimensional system
\begin{align}
    \dot{k} &= f(k) - c - \delta k, \\
    \frac{\dot{c}}{c} &= \frac{1}{\sigma}\bigl(f'(k) - \delta - \rho\bigr).
\end{align}


\subsection{Solving a System of ODEs}

Consider a linear (or linearised) continuous-time system of the form
\begin{equation}
    \dot{\mathbf{z}}(t) = A\,\mathbf{z}(t),
\end{equation}
where $\mathbf{z}$ collects deviations of the variables from their steady-state values.
The solution procedure has four steps.

\begin{enumerate}
    \item \textbf{Find the eigenvalues.} Solve $\det(A - \mu I) = 0$ for the eigenvalues
          $\mu_1, \mu_2$ of $A$.

    \item \textbf{Find the eigenvectors.} For each $\mu_i$, solve
          $(A - \mu_i I)\mathbf{v}_i = \mathbf{0}$ for the eigenvector $\mathbf{v}_i$.

    \item \textbf{Write the general solution.}
          \begin{equation}
              \mathbf{z}(t) = \alpha_1\, e^{\mu_1 t}\,\mathbf{v}_1
                             + \alpha_2\, e^{\mu_2 t}\,\mathbf{v}_2.
          \end{equation}

    \item \textbf{Apply boundary conditions.} Use the initial condition $\mathbf{z}(0)$ and
          any terminal/transversality condition to determine $\alpha_1, \alpha_2$.
\end{enumerate}

\paragraph{Stability.}
The nature of the eigenvalues determines local dynamics:
\begin{itemize}
    \item If all eigenvalues have \emph{negative} real parts, the steady state is
          \textbf{locally stable} (a node or spiral).
    \item If the eigenvalues have \emph{opposite} signs ($\mu_1 < 0 < \mu_2$), the steady
          state is a \textbf{saddle point}.
\end{itemize}

\paragraph{Saddle-path stability in economic models.}
In optimisation-based macro models, the state variable $k$ is predetermined (initial
condition given), while the control variable $c$ is a free ``jump'' variable. The
optimal solution therefore requires setting the coefficient on the unstable
eigenvector to zero ($\alpha_2 = 0$, corresponding to $\mu_2 > 0$). This selects the
unique \textbf{saddle path} — the only trajectory compatible with the TVC — and pins
down the correct initial value $c_0$.


\subsection{random}

inelastically labor supply means a vertival line in a w - L plot (same L for any w). -> this applies to ramsey, solow, romer


    \item \textit{Stock vs.\ flow:} A stock is measured at a given point in time; a flow is measured over a period of time. GDP is a flow (``GDP in 2024 was \dots''); the amount of money in your pocket right now is a stock.
    \item \textit{Side note on labor productivity:} \textit{Average} productivity of labor is $Y/L$, while \textit{marginal} productivity of labor is $\partial Y/\partial L$.